---
title: "Customer Churn"
author: "Madan K C"
date: "2023-04-27"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Customer Churn Prediction
In this project, we will be analyzing a telecom churn dataset. The objective of the analysis is to develop a model to predict whether a customer will churn or not. The dataset consists of 7043 observations and 23 variables. The dataset contains a mix of categorical and continuous variables.
I have copied the dataset from kaggle:https://www.kaggle.com/datasets/datazng/telecom-company-churn-rate-call-center-data/discussion/392565


Load data
```{r}
library(readxl)
telecom_data <- read_excel("data/telecom-churn-rate-dataset.xlsx")
#View(telecom_churn_rate_dataset)
```

```{r}
dim(telecom_data)
```

```{r}
names(telecom_data)
```


```{r}
str(telecom_data)
```

Find the number of unique values in each variable.
```{r}
library(dplyr)
sapply(telecom_data, n_distinct)
```


There are not any missing values in the dataset.
```{r}
colSums(is.na(telecom_data))
```


Remove the observations having missing values.
```{r}
telecom_data <- na.omit(telecom_data)
```


```{r}
dim(telecom_data)
```

Now the rows having missing values have been removed. 
Checking if there are any mising values again.
```{r}
colSums(is.na(telecom_data))
```


Converting the categorical variables to factor.
```{r}
telecom_data[, sapply(telecom_data, is.character)] <- lapply(telecom_data[, sapply(telecom_data, is.character)], factor)

```

Check whether the categorical are now factor or not.
```{r}
str(telecom_data)
```


```{r}
library(magrittr)
library(dplyr)
customer_status <- telecom_data %>% group_by(gender) %>% summarise(num_customers = n())
customer_status
```

As Customer ID is not related to the regression, using variables except customerID variable.
Removing this from the dataframe
```{r}
telecom_data <- select(telecom_data, -customerID)
```


Logistic regression with all other variables considered.
```{r}
logistic_reg1 <- glm(Churn ~ ., family = binomial, data = telecom_data)
summary(logistic_reg1)
```


Considering only the variables having significant

modeling_data = telecom_data %>% dplyr::select(-customerID)


```{r}
logistic_reg <- glm(Churn ~ SeniorCitizen + tenure + MultipleLines + Contract + PaperlessBilling + TotalCharges + numTechTickets, family = binomial, data = telecom_data)
summary(logistic_reg)
```


## Data Partition and Modelling
```{r}
library(caret)
data_partition <- telecom_data
#telecom_data$Churn = as.factor(telecom_data$Churn)
data_partition$leave = ifelse(data_partition$Churn == "Yes","Churn","Not")
data_partition$leave = as.factor(data_partition$leave)
data_partition = data_partition %>% dplyr::select(-Churn) #removing the column with numbers, otherwise the prediction is obvious

set.seed(1)
test.indices = createDataPartition(data_partition$leave, p = 0.2, list = FALSE) #classic 80/20 train-test partition
test_partition = data_partition[test.indices,]
train_partition = data_partition[-test.indices,]
```



```{r}
model_train = train(leave ~ SeniorCitizen + tenure + MultipleLines + Contract + PaperlessBilling + TotalCharges + numTechTickets, data = train_partition, method = "glm", family = binomial(link = "logit"))

predTrain = predict(model_train, train_partition)
predTest = predict(model_train, test_partition)
```

For Training data, the Confusion Matrix:
```{r}
confusionMatrix(predTrain, train_partition$leave, positive = "Churn")
```
For Testing data, the Confusion Matrix:
```{r}
confusionMatrix(predTest, test_partition$leave, positive = "Churn")

```

## Cross-validation

```{r}
train_control <- trainControl(method="cv", number=15) #10-fold cross validation
model_cv <- caret::train(leave ~ SeniorCitizen + tenure + MultipleLines + Contract + PaperlessBilling + TotalCharges + numTechTickets, data=train_partition, trControl=train_control, method = "glm", family = binomial(link = "logit"))
print(model_cv)
```

```{r}
predTest.cv <- predict(model_cv, test_partition)
cmTest.cv = confusionMatrix(predTest.cv, test_partition$leave)
cmTest.cv
```

```{r}
importance <- varImp(model_train, scale=FALSE)
plot(importance)
```

## Decision Tree

```{r}
library(party)
tree <- ctree(leave~., data = train_partition)
```

```{r}
treePredTrain <- predict(tree, train_partition, type = "response")

confusionMatrix(treePredTrain,train_partition$leave)
```

```{r}
treePredTest <- predict(tree, test_partition, type = "response")

confusionMatrix(treePredTest,test_partition$leave)
```


Cross Validation: Decision Tree

```{r}
train_control_tree <- trainControl(method="cv", number=10)
model_tree <- caret::train(leave~., data=train_partition, trControl=train_control_tree, method="ctree")

print(model_tree)
```

```{r}
predTest_tree <- predict(model_tree, test_partition)
tree_cv = confusionMatrix(predTest_tree, test_partition$leave)
tree_cv

```




